### 2018-01-08 manuscript-md/06_figures_tables.Rmd
Isolate code from Rmd in separate file
This file will appear at the end of the manuscript, following the
formatting requirements of New Phytologist.

Ideally, I'd like to show only figure captions, but not the figures
themselves, so that the figure labels are correctly included in the
text.

<!-- Methods -->

```{r "pfts"}
n_species <- try_data %>%
  distinct(clm45, AccSpeciesID) %>%
  select(pft = clm45, species = AccSpeciesID) %>%
  count(pft) %>%
  mutate(
    pft = as.character(pft),
    label = pft2abbr[pft],
    pft = tools::toTitleCase(gsub("_", " ", pft))
  ) %>%
  select(`Label` = label, `PFT` = pft, `Number of species` = n)
species_caption <- paste(
  "Names, labels, and species counts for plant functional types (PFTs)",
  "used in this analysis."
)
kable(n_species, caption = species_caption, align = rep("c", ncol(n_species)))
```

(ref:ssfigcap) Sample sizes for each trait-PFT pair. $y$ axis is scaled logarithmically.

```{r "samplesize", fig.cap='(ref:ssfigcap)', fig.width=7, fig.height=7}
ss_plot <- sample_size %>%
  mutate(param = forcats::lvls_revalue(param, param_simple[both_params])) %>%
  ggplot() + 
  aes(x = pft, y = sample_size, fill = pft) + 
  geom_col() +
  scale_y_log10(breaks = c(10, 100, 1000, 10000)) +
  facet_wrap(~param, ncol = 2, labeller = label_parsed) +
  scale_fill_manual(values = pft_colors) +
  xlab("Plant functional type") +
  ylab("Sample size") +
  guides(fill = guide_legend(title = "PFT")) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
ss_plot
```

<!-- Results -->

```{r prepresults}
cachefile <- ".cache/mvtraits_results.rds"
results_all <- readRDS(cachefile)

max_ss <- max(sample_size$sample_size)

multi_summary <- results_all %>%
  filter(model_type %in% c("uni", "multi")) %>%
  mutate(
    mu_mean = map(data, c("mu", "Mean")) %>%
      map(~as_tibble(as.list(.))),
    mu_lo = map(data, c("mu", "2.5%")) %>%
      map(~as_tibble(as.list(.))),
    mu_hi = map(data, c("mu", "97.5%")) %>%
      map(~as_tibble(as.list(.)))
      ) %>%
  select(model_type:pft, mu_mean:mu_hi) %>%
  unnest(.sep = ".") %>%
  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
  separate(variable, c("stat", "param"), sep = "\\.") %>%
  spread(stat, value) %>%
  mutate(
    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr),
    param = factor(param, both_params)
  )

hier_prep <- results_all %>%
  filter(model_type == "hier") %>%
  select(model_type:pft_type, data) %>%
  mutate(
    mu_vals = map(data, "mu_group"),
    pft = map(mu_vals, names),
    mu_df = map2(pft, mu_vals, ~tibble(pft = ..1, dat = ..2))
  ) %>%
  select(-data, -mu_vals, -pft) %>%
  unnest()

hier_summary <- hier_prep %>%
  mutate(
    mu_mean = map(dat, c("Mean")) %>%
      map(~as_tibble(as.list(.))),
    mu_lo = map(dat, c("2.5%")) %>%
      map(~as_tibble(as.list(.))),
    mu_hi = map(dat, c("97.5%")) %>%
      map(~as_tibble(as.list(.)))
      ) %>%
  select(model_type:pft, mu_mean:mu_hi) %>%
  unnest(.sep = ".") %>%
  gather("variable", "value", -(model_type:pft), na.rm = TRUE) %>%
  separate(variable, c("stat", "param"), sep = "\\.") %>%
  spread(stat, value) %>%
  mutate(
    pft = if_else(is.na(pft), "global", pft),
    pft = factor(pft, abbr2pft) %>% forcats::lvls_revalue(pft2abbr),
    param = factor(param, both_params)
  )

# Absolute comparison of means
means_dat <- multi_summary %>%
  filter(pft != "GLOB", !is.na(pft)) %>%
  bind_rows(hier_summary) %>%
  left_join(sample_size) %>%
  mutate(
    model_type = factor(model_type, c("uni", "multi", "hier")) %>%
      forcats::lvls_revalue(c("univariate", "multivariate", "hierarchical")),
    mass_area = factor(mass_area, c("mass", "area")),
  ) %>%
  mutate_at(
    c("mu_mean", "mu_lo", "mu_hi"),
    ~10 ^ .
  ) %>%
  filter(
    !(param %in% c("leaf_lifespan", "SLA") & mass_area == "area")
  )
```

```{r meanvaluetable, results = "asis"}
## Write summary table
table_cap_mass <- paste(
  "Mean and 95\\% confidence interval of trait estimates",
  "for mass-normalized traits from the hierarchical model."
)
table_cap_area <- gsub("mass", "area", table_cap_mass)

censor <- function(rxp, pft, pfts, param, value) {
  if_else(grepl(rxp, param) & pfts %in% pft, "---", value)
}

table_dat <- means_dat %>%
  filter(
    model_type == "hierarchical",
    !(param %in% c("SLA", "leaf_lifespan") & mass_area == "area")
  ) %>%
  mutate(
    mu_lo = if_else(param == "Rdmass", mu_lo * 1000, mu_lo),
    mu_mean = if_else(param == "Rdmass", mu_mean * 1000, mu_mean),
    mu_hi = if_else(param == "Rdmass", mu_hi * 1000, mu_hi),
    mu_lo = if_else(param == "Rdarea", mu_lo * 1000, mu_lo),
    mu_mean = if_else(param == "Rdarea", mu_mean * 1000, mu_mean),
    mu_hi = if_else(param == "Rdarea", mu_hi * 1000, mu_hi),
    lo_f = formatC(mu_lo, digits = 3),
    mid_f = formatC(mu_mean, digits = 3),
    hi_f = formatC(mu_hi, digits = 3),
    value = sprintf("%s (%s,%s)", mid_f, lo_f, hi_f),
    value = censor("Vcmax|Jmax", "C4G", pft, param, value),
    value = censor("Rdmass", "ShDBo", pft, param, value),
    value = censor("Jmax_mass", c("NlEBo", "NlD"), pft, param, value),
    value = censor("Jmax_area", c("NlD"), pft, param, value)
  ) %>%
  select(pft, param, value) %>%
  mutate(
    param = factor(param, both_params) %>% 
      forcats::lvls_revalue(param_markdown_nounit[both_params]),
  )%>%
  spread(param, value) %>%
  arrange(pft)

table_dat %>%
  select(-matches("area")) %>%
  kable(caption = table_cap_mass, format = "latex",
        escape = FALSE, booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("scale_down"))
table_dat %>%
  select(-matches("mass")) %>%
  kable(caption = table_cap_area, format = "latex",
        escape = FALSE, booktabs = TRUE) %>%
  kableExtra::kable_styling(latex_options = c("scale_down"))
```

(ref:mccap) Mean and 95% confidence interval on best estimates of traits for each plant functional type from the univariate, multivariate, and hierarchical models.
For leaf lifespan and SLA, results were not significantly different between the mass- and area-based models, so only results from the mass-based model are shown.
For some PFT-trait combinations, where large error bars resulting from the relatively uninformative priors are substantially larger than the variability among means, the $y$ axes are constrained to facilitate comparison.

```{r meancompare, fig.width=7, fig.height=7, fig.cap='(ref:mccap)'}
pclip <- function(p, pname, value, lim, hi = TRUE) {
  f <- ifelse(hi, `>`, `<`)
  if_else(p == pname & f(value, lim), lim, value)
}

plot_dat <- means_dat %>%
  mutate(
    #too_few = sample_size < 2,
    #mu_hi = if_else(too_few, NA_real_, mu_hi),
    #mu_lo = if_else(too_few, NA_real_, mu_lo),
    #mu_mean = if_else(too_few, NA_real_, mu_mean),
    irrelevant = grepl("Vcmax|Jmax", param) & pft == "C4G",
    mu_hi = if_else(irrelevant, NA_real_, mu_hi),
    mu_lo = if_else(irrelevant, NA_real_, mu_lo),
    mu_mean = if_else(irrelevant, NA_real_, mu_mean),
    ## TODO: Fix these in the prior and re-run
    mu_hi = pclip(param, "Vcmax_mass", mu_hi, 1.5),
    mu_hi = pclip(param, "Vcmax_area", mu_hi, 100),
    mu_hi = pclip(param, "Jmax_mass", mu_hi, 1.8),
    mu_hi = pclip(param, "Jmax_area", mu_hi, 150),
    mu_mean = pclip(param, "Jmax_area", mu_mean, 150),
    mu_hi = pclip(param, "Rdmass", mu_hi, 0.03),
    mu_mean = pclip(param, "Rdmass", mu_mean, 0.03),
    mu_lo = pclip(param, "Rdmass", mu_lo, 0.03),
    mu_hi = pclip(param, "Rdarea", mu_hi, 0.0025),
    mu_mean = pclip(param, "Rdarea", mu_mean, 0.0025),
    mu_lo = pclip(param, "Rdarea", mu_lo, 0.0025),
    mu_hi = pclip(param, "Parea", mu_hi, 0.285),
    #mu_lo = pclip(param, "Jmax_area", mu_lo, 40, FALSE),
    param = forcats::lvls_revalue(param, param_fancy_chr[both_params])
  )

p_means <- ggplot(plot_dat) +
  aes(x = interaction(model_type, pft),
      y = mu_mean, ymin = mu_lo, ymax = mu_hi,
      color = pft, shape = model_type) +
  geom_pointrange(size = 0.3) +
  facet_wrap(~param, scales = "free", ncol = 2,
              labeller = label_parsed) +
  scale_y_continuous() +
  scale_color_manual(values = pft_colors) +
  guides(
    shape = guide_legend(title = "Model type"),
    color = guide_legend(title = "PFT")
  ) +
  ylab("Trait estimate mean and 95% CI") +
  xlab("Model type and PFT") +
  theme_bw() +
  theme(
    text = element_text(size = 12),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank()
  )
p_means
```

```{r "civssamplesize", fig.width=4.4, fig.height=3, fig.cap='(ref:sscap)'}
## Relative uncertainty
cv_dat <- means_dat %>%
  mutate(
    cv = abs((mu_hi - mu_lo) / mu_mean)
  ) %>%
  filter(sample_size > 1)

cv_mod <- cv_dat %>%
  group_by(model_type) %>%
  nest() %>%
  mutate(
    cvfit = map(data, ~lm(log10(cv) ~ log10(sample_size), data = .))
  )

cv_coef <- cv_mod %>%
  mutate(
    coef = map(cvfit, broom::tidy)
  ) %>%
  unnest(coef) %>%
  select(model_type, term, estimate) %>%
  spread(term, estimate) %>%
  "names<-"(c("model_type", "intercept", "slope")) %>%
  mutate(
    eqn = sprintf("'%s: '~log[10](y) == %0.2f - %0.2f~log[10](x)", model_type, intercept, abs(slope))
  )

cv_pred <- cv_mod %>%
  mutate(
    grid = map(data, modelr::data_grid, sample_size = seq(5, max_ss)),
    cvpred = map2(grid, cvfit, modelr::add_predictions)
  ) %>%
  unnest(cvpred) %>%
  mutate(cv = 10 ^ pred)

x_lims <- range(cv_pred$sample_size)
y_lims <- range(cv_pred$cv)

relative_ci <- ggplot(cv_dat) +
  aes(x = sample_size, y = cv, color = model_type) +
  geom_point(size = 0.5) +
  geom_line(data = cv_pred) +
  annotate("text", x = rep(2500, 3), y = c(1.4, 1.25, 1.1), 
           label = cv_coef$eqn, parse = TRUE, 
           size = 2, vjust = "right") +
  scale_x_continuous(trans = "log10") +
  coord_cartesian(xlim = x_lims, ylim = y_lims) +
  xlab("Sample size") +
  ylab("Relative width of 95% CI") +
  scale_color_brewer(type = "qual", palette = 2) +
  guides(color = guide_legend(title = "Model type")) +
  theme_bw() +
  theme(
    legend.position = c(0.85, 0.75),
    panel.grid = element_blank()
  )
relative_ci
```

(ref:stickcap) Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
Covariance estimates not significantly different from zero ($p < 0.05$) are indicated by x symbols at the mean estimate.
$x$ and $y$ axes vary on a log scale, reflecting the fact that the model was fit using the base 10 log of all traits.
With the exception of leaf lifespan, pairwise covariances are consistent in direction but vary somewhat in magnitude between PFTs, and when comparing PFT-level and global estimates.
However, many pairwise covariances are not statistically significant, particularly (but not always) for undersampled traits and PFTs.

```{r stickpairs, fig.width = 5, fig.height = 5, fig.cap = '(ref:stickcap)'}
suppressPackageStartupMessages({
  library(mvtraits)
})
results_sub <- results_all %>%
  filter((model_type == "multi" & pft == "global") |
         (model_type == "hier"))
dat_list <- results_sub$data
names(dat_list) <- with(results_sub, paste(model_type, mass_area, sep = "_"))

stickplot_pairs(
  #mu_global_lower = dat_list$hier_mass$mu_global,
  #Sigma_global_lower = dat_list$hier_mass$Sigma_global,
  mu_global_lower = dat_list$multi_mass$mu,
  Sigma_global_lower = dat_list$multi_mass$Sigma,
  mu_group_lower = dat_list$hier_mass$mu_group,
  Sigma_group_lower = dat_list$hier_mass$Sigma_group,
  vars_lower = mass_params,
  group_names = pft2abbr[-1],
  vars_label = param_labels,
  unlog_axes = TRUE,
  #mu_global_upper = dat_list$hier_area$mu_global,
  #Sigma_global_upper = dat_list$hier_area$Sigma_global,
  mu_global_upper = dat_list$multi_area$mu,
  Sigma_global_upper = dat_list$multi_area$Sigma,
  mu_group_upper = dat_list$hier_area$mu_group,
  Sigma_group_upper = dat_list$hier_area$Sigma_group,
  vars_upper = area_params,
  col_group = pft_colors,
  cex_g = 1.2,
  lwd_g = 1,
  lty_g = 1,
  lwd_s = 1.3,
  par_plot = list(oma = c(0, 5, 3, 5), cex = 0.4),
  par_legplot = list(mar = c(0, 0, 0, 0)),
  par_legend = list(ncol = 5, x = "center", cex = 0.6),
  reorder_legend = TRUE
)
mtext("Mass-normalized", side = 2, cex = 0.7, line = 3)
mtext("Area-normalized", side = 4, cex = 0.7, line = 1)
```

```{r boxprep}
outdir <- "../output"
hier_mass_file <- tail(dir(outdir, "hier.mass.clm45", full.names = TRUE), 1)
hier_area_file <- tail(dir(outdir, "hier.area.clm45", full.names = TRUE), 1)
multi_mass_file <- tail(dir(outdir, "multi.mass.clm45", full.names = TRUE), 1)
multi_area_file <- tail(dir(outdir, "multi.area.clm45", full.names = TRUE), 1)

hier_summary_mass <- readRDS(hier_mass_file)[["summary_table"]] %>%
  mutate(mass_area = "mass")
hier_summary_area <- readRDS(hier_area_file)[["summary_table"]] %>%
  mutate(mass_area = "area")
multi_summary_mass <- readRDS(multi_mass_file)[["summary_table"]] %>%
  mutate(mass_area = "mass")
multi_summary_area <- readRDS(multi_area_file)[["summary_table"]] %>%
  mutate(mass_area = "area")

proc <- function(dat, hier = TRUE) {
  if (hier) {
    dat <- filter(dat, group != "global")
  } else {
    dat <- mutate(dat, group = "global")
  }
  dat <- dat %>%
    filter(variable == "Corr") %>%
    separate(index, c("xvar", "yvar"), sep = "\\.\\.") %>%
    mutate(
      pft = factor(group, abbr2pft) %>% forcats::lvls_revalue(pft2abbr)
    ) %>%
    select(mass_area, pft, xvar, yvar, Mean, `2.5%`, `97.5%`)
}

hsm <- proc(hier_summary_mass)
hsa <- proc(hier_summary_area)
msm <- proc(multi_summary_mass, FALSE)
msa <- proc(multi_summary_area, FALSE)
hs <- bind_rows(hsm, hsa, msm, msa)
```

(ref:corrbarscap) Mean and 95% CI on estimates of pairwise correlation coefficients
for all data pooled globally (dark grey) and for each PFT (colored).
For most PFT-trait pairs, correlations are mutually consistent in magnitude but vary in strength.

```{r corrbars, fig.width=7, fig.height=7, fig.cap='(ref:corrbarscap)'}
.zzz <- close.screen(all.screens = TRUE)
corr_bar <- function(dat, xvar, yvar, mass_area) {
  dp <- dat %>%
    filter(
      mass_area == !!mass_area,
      (xvar == !!xvar & yvar == !!yvar) | 
        (xvar == !!yvar & yvar == !!xvar),
    ) %>%
    arrange(pft) %>%
    mutate(color = c("grey25", pft_colors))
  x <- barplot(dp$Mean, col = dp$color, border = dp$color,
               ylim = c(-1, 1), axes = FALSE)
  arrows(x, dp[["2.5%"]], x, dp[["97.5%"]],
         length = 0.01, angle = 90, code = 3)
  box()
}

nv <- length(mass_params)
main_screens <- split.screen(rbind(c(0, 1, 0.2, 1), c(0, 1, 0, 0.2)))
par(mar = rep(0.1, 4), oma = c(0, 8, 1, 5), cex = 0.4)
screens <- split.screen(c(nv, nv), screen = main_screens[1])
for (i in seq_len(nv)) {
  for (j in seq_len(nv)) {
    k <- j + nv * (i - 1)
    screen(screens[k])
    if (i == j) {
      # Label
      plot(0:1, 0:1, type = "n", ann = FALSE, axes = FALSE)
      box()
      text(x = 0.5, y = 0.5, labels = param_labels[i], cex = 2)
    } else {
      if (j > i) {
        # Upper triangle
        mass_area <- "area"
        xvar <- area_params[j]
        yvar <- area_params[i]
      } else {
        # Lower triangle
        mass_area <- "mass"
        xvar <- mass_params[j]
        yvar <- mass_params[i]
      }
      corr_bar(hs, xvar, yvar, mass_area)
    }
    if (i == 2 && j == 1) {
      axis(side = 2)
    }
    if (i == 1 && j == nv) {
      axis(side = 4)
    }
  }
}
## Draw legend
screen(main_screens[2])
par(mar = c(0, 0, 0, 0), oma = c(0, 0, 0, 0))
plot(0, 0, type = "n", axes = FALSE, ann = FALSE)
legend("center", legend = levels(hs$pft), fill = c("grey25", pft_colors), 
       ncol = 5, bty = "n")
close.screen(all.screens = TRUE)
mtext("Mass-normalized", side = 2, cex = 0.7, line = 1)
mtext("Mean and 95% CI of pairwise correlation", side = 2, cex = 1, line = 2)
mtext("Area-normalized", side = 4, cex = 0.7, line = 0)
```