### 2018-02-13 manuscript/06_figures_tables.Rmd
Revise manuscript build process
- Use `doc` as default output
- Remove figures (draw them separately; only use figure captions in
manuscript)
- Render tables using `knitr`, but at end of manuscript

# Tables {-}

```{r "knitr_options", echo = FALSE, message = FALSE, warning = FALSE, error = FALSE}
library(knitr)
library(here)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  error = FALSE
)
options(knitr.table.format = "pandoc")
```

<!-- Tables ---------------------------->
```{r "read_chunks"}
knitr::read_chunk(here::here("scripts/manuscript-analyses/try_ref_table.R"))
knitr::read_chunk(here::here("scripts/manuscript-analyses/table_n_species.R"))
```

<!-- TRY data citations -->
```{r "trydata"}
```

<!-- Number of species in each PFT -->
```{r "pfts"}
```

# Figure captions {-}

**Figure 1**: Sample sizes for each trait-PFT pair. $y$ axis is scaled logarithmically.

**Figure 2**: Mean and 95% confidence interval on best estimates of traits for each plant functional type from the univariate, multivariate, and hierarchical models.
For leaf lifespan and SLA, results were not significantly different between the mass- and area-based models, so only results from the mass-based model are shown.
For some PFT-trait combinations, where large error bars resulting from the relatively uninformative priors are substantially larger than the variability among means, the $y$ axes are constrained to facilitate comparison.

**Figure 3**: Relative uncertainty in PFT-level trait estimates as a function of sample size for each model type.
Lines represent linear models ($\log(y) = b_0 + b_1 \log(x)$) fit independently for each model type.
In general, differences in estimate uncertainty between the univariate and multivariate models were minimal at large sample sizes but increasingly important at low sample sizes.
However, differences in estimate uncertainty between the multivariate and hierarchical models were consistently negligible.

**Figure 4**: Pairwise trait mean and covariance estimates for all data pooled globally (black) and for each PFT (colored).
Covariance estimates not significantly different from zero ($p < 0.05$) are indicated by x symbols at the mean estimate.
$x$ and $y$ axes vary on a log scale, reflecting the fact that the model was fit using the base 10 log of all traits.
With the exception of leaf lifespan, pairwise covariances are consistent in direction but vary somewhat in magnitude between PFTs, and when comparing PFT-level and global estimates.
However, many pairwise covariances are not statistically significant, particularly (but not always) for undersampled traits and PFTs.

**Figure 5**: Mean and 95% CI on estimates of pairwise correlation coefficients
for all data pooled globally (dark grey) and for each PFT (colored).
For most PFT-trait pairs, correlations are mutually consistent in magnitude but vary in strength.

# References {-}