### 2018-06-18 supplementary_information.Rmd
Add data range analysis and some more review responses

---
title: Supplementary information
author:
  - >
    Alexey N. Shiklomanov,
    Elizabeth M. Cowdery,
    Michael Bahn,
    Chaeho Byun,
    Joseph Craine,
    Andrés Gonzalez-Melo,
    Steven Jansen,
    Nathan Kraft,
    Koen Kramer,
    Vanessa Minden,
    Ülo Niinemets,
    Yusuke Onoda,
    Enio Egon Sosinski,
    Nadejda A. Soudzilovskaia,
    Michael C. Dietze
output: pdf_document
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
---

```{r libraries, include = FALSE}
library(knitr)
library(kableExtra)
library(drake)
library(ggplot2)

correlation_scatter_cap <- paste(
  "Pairwise mean squared trait correlation values as a function of sample size.",
  "Note that the x-axis uses a $\\log_{10}$ scale."
)

correlation_fig_cap <- paste(
  "Fraction of statistically significant ($p < 0.05$) pairwise trait correlation estimates",
  "as a function of sample size for each plant functional type."
)

correlation_range_cap <- paste(
  "Relationship between pairwise correlation strength and data range for each PFT.",
  "In each pair of sub-panels, the left panel is relationship between pairwise correlation",
  "and the range of the variable on the x axis",
  "while the right panel is the relationship between pairwise correlation",
  "and the y variable",
  "(For instance, in the two-panel figure in row 1 column 2,",
  "the left panel is the SLA-leaf lifespan correlation as a function of the range of SLA",
  "and the right panel is the same correlation as a function of the range of leaf lifespan).",
  "Data range is expressed as the difference between the 2.5% and 97.5% quantiles of each normalized trait (i.e. $\\frac{X}{E[X]}$).",
  "Grey shading is the 95% confidence interval around a local regression (LOESS).",
  "Upper triangle is area-normalized values, and lower triangle is mass-normalized values."
)
```

```{r mean_value_table, echo = FALSE, results = "asis"}
vals <- readd()
```

```{r sample_size_table, echo = FALSE, results = "asis"}
ssw <- readd(sample_size_wide)
kable(
  ssw,
  col.names = c(
    "PFT", "Leaf lifespan", "SLA",
    "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
    "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
    "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
    "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
    "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
  ),
  format = "latex",
  booktabs = TRUE,
  escape = FALSE,
  caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
) %>%
  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
  group_rows("Broadleaf", 2, 6) %>%
  group_rows("Needleleaf", 7, 9) %>%
  group_rows("Shrub", 10, 12) %>%
  group_rows("Grass", 13, 15) %>%
  landscape()
```

```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
corr_results <- readd(corr_processed)
kable(
  corr_results %>%
    dplyr::select(pft, yvar, xvar, corr_value, present, missing),
  "latex",
  booktabs = TRUE, longtable = TRUE, escape = FALSE,
  col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95\\% CI)", "present", "missing"),
  caption = "Pairwise trait correlation values for each plant functional type. Values with asterisks indicate correlations significantly different from zero ($p < 0.05$)."
)
```

```{r correlation_sample_size_pairs, echo = FALSE, fig.cap = correlation_scatter_cap}
corr_results %>%
  mutate(
    significant = factor(significant, c("TRUE", "FALSE")) %>%
      forcats::lvls_revalue(c("yes", "no"))
  ) %>%
  ggplot() +
  aes(x = present + 0.1, y = Mean, color = significant) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~ pft) +
  scale_x_log10() +
  labs(x = "Sample size", y = "Correlation") +
  theme_bw() +
  theme(
	legend.position = "bottom"
  )
```

```{r correlation_sample_size_analysis, echo = FALSE, include = FALSE, eval = FALSE, fig.cap = correlation_fig_cap}
corr_binned <- corr_results %>%
  dplyr::mutate(
    ss_bin = cut(
      present,
      c(0, 1, 10, 25, 50, 100, 500, 1000, Inf),
      c("0", "< 10", "< 25", "< 50", "< 100", "< 500", "< 1000", "> 1000"),
      include.lowest = TRUE,
      right = FALSE
    )
  ) %>%
  group_by(pft, ss_bin) %>%
  summarize(frac_significant = mean(significant))

ggplot(corr_binned) +
  aes(x = ss_bin, y = frac_significant) +
  geom_segment(aes(xend = ss_bin, yend = 0)) +
  geom_point() +
  facet_wrap(~ pft) +
  labs(x = "Sample size", y = "Fraction of significant correlations") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.9))
```

```{r correlation_range_analysis, echo = FALSE, results = "hide", warning = FALSE, fig.cap = correlation_range_cap, fig.dim = c(12, 12)}
readd(corr_range_plot_gg)
```