### 2018-09-07 manuscript/05_figures_tables.Rmd

<div custom-style="Heading, no number">Tables</div>

```{r "knitr_options", echo = FALSE, message = FALSE, warning = FALSE, error = FALSE}
library(knitr)
library(here)
opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE, error = FALSE
)
options(knitr.table.format = "pandoc")
```

<!-- Tables ---------------------------->
```{r "read_chunks"}
knitr::read_chunk(here::here("scripts/manuscript-analyses/try_ref_table.R"))
knitr::read_chunk(here::here("scripts/manuscript-analyses/table_n_species.R"))
```

<!-- TRY data citations -->
```{r "trydata"}
```

<!-- Number of species in each PFT -->
```{r "pfts"}
```