### 2019-09-03 supplementary_information.Rmd
Janky but functional code for RMA table

---
title: Supporting information
author:
  - >
    Alexey N. Shiklomanov,
    Elizabeth M. Cowdery,
    Michael Bahn,
    Chaeho Byun,
    Joseph Craine,
    Andrés Gonzalez-Melo,
    Steven Jansen,
    Koen Kramer,
    Vanessa Minden,
    Ülo Niinemets,
    Yusuke Onoda,
    Enio Egon Sosinski,
    Nadejda A. Soudzilovskaia,
    Michael C. Dietze
output: pdf_document
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage[table]{xcolor}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \renewcommand{\thetable}{S\arabic{table}}
  - \renewcommand{\thefigure}{S\arabic{figure}}
---

```{r libraries, include = FALSE}
library(knitr)
library(kableExtra)
## library(drake)
library(dplyr, exclude = "group_rows")
library(tidyverse)

correlation_fig_cap <- paste(
  "Fraction of statistically significant ($p < 0.05$) pairwise trait correlation estimates",
  "as a function of sample size for each plant functional type."
)
```

```{r mean_value_table, echo = FALSE, results = "asis"}
table_cap_mass <- paste(
  "Mean and 95\\% confidence interval of trait estimates",
  "for mass-normalized traits from the hierarchical model.",
  "Note that some traits for some PFTs have effectively no constraint relative to the mostly uninformative prior (as evidenced by their very large confidence intervals),",
  "so their values should be used with caution."
)

table_cap_area <- paste(
  "Mean and 95\\% confidence interval of trait estimates",
  "for area-normalized traits from the hierarchical model.",
  "Note that some traits for some PFTs have effectively no constraint relative to the mostly uninformative prior (as evidenced by their very large confidence intervals),",
  "so their values should be used with caution.",
  "See mass-normalized table for SLA and leaf lifespan estimates."
)

with_clm <- readRDS(here::here("extdata", "tidy-means.rds"))

# Mass-normalized table
with_clm %>%
  filter(mass_area == "mass",
         model_type == "hierarchical") %>%
  select(pft, param, mu_hi, mu_lo, mu_mean) %>%
  # Fix units
  mutate_at(
    vars(starts_with("mu")),
    ~if_else(param %in% c("Nmass", "Pmass"), . * 1000, .)
  ) %>%
  mutate(string = sprintf("%.2f (%.2f, %.2f)",
                          mu_mean, mu_lo, mu_hi)) %>%
  select(pft, param, string) %>%
  spread(param, string) %>%
  arrange(pft) %>%
  kable(
    caption = table_cap_mass,
    format = "latex",
    col.names = linebreak(c(
      "PFT",
      "Leaf lifespan\n(months)",
      "SLA\n(m$^2$ kg$^{-1}$)",
      "$N_\\textrm{mass}$\n(mg g$^{-1}$)",
      "$P_\\textrm{mass}$\n(mg g$^{-1}$)",
      "$R_\\textrm{d, mass}$\n($\\mu$mol g$^{-1}$ s$^{-1}$)",
      "$V_\\textrm{c, max, mass}$\n($\\mu$mol g$^{-1}$ s$^{-1}$)",
      "$J_\\textrm{max, mass}$\n($\\mu$mol g$^{-1}$ s$^{-1}$)"
    )),
    escape = FALSE,
    booktabs = TRUE
  ) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))

with_clm %>%
  filter(mass_area == "area",
         model_type == "hierarchical") %>%
  select(pft, param, mu_hi, mu_lo, mu_mean) %>%
  mutate(string = sprintf("%.2f (%.2f, %.2f)",
                          mu_mean, mu_lo, mu_hi)) %>%
  select(pft, param, string) %>%
  spread(param, string) %>%
  arrange(pft) %>%
  kable(
	caption = table_cap_area,
	format = "latex",
    col.names = linebreak(c(
      "PFT",
      ## "Leaf lifespan\n(months)",
      ## "SLA\n(m$^2$ kg$^{-1}$)",
      "$N_\\textrm{area}$\n(g m$^{-2}$)",
      "$P_\\textrm{area}$\n(g m$^{-2}$)",
      "$R_\\textrm{d, area}$\n($\\mu$mol m$^{-2}$ s$^{-1}$)",
      "$V_\\textrm{c, max, area}$\n($\\mu$mol m$^{-2}$ s$^{-1}$)",
      "$J_\\textrm{max, area}$\n($\\mu$mol m$^{-2}$ s$^{-1}$)"
    )),
	escape = FALSE,
	booktabs = TRUE
) %>%
  kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))

```

```{r sample_size_table, echo = FALSE, results = "asis"}
## ssw <- readd(sample_size_wide)
## kable(
##   ssw %>%
##     dplyr::select(pft, leaf_lifespan, SLA,
##                   Nmass, Narea, Pmass, Parea, Rdmass, Rdarea,
##                   Vcmax_mass, Vcmax_area, Jmax_mass, Jmax_area),
##   col.names = c(
##     "PFT", "Leaf lifespan", "SLA",
##     "$N_\\textrm{mass}$", "$N_\\textrm{area}$",
##     "$P_\\textrm{mass}$", "$P_\\textrm{area}$",
##     "$R_\\textrm{d, mass}$", "$R_\\textrm{d, area}$",
##     "$V_\\textrm{c, max, mass}$", "$V_\\textrm{c, max, area}$",
##     "$J_\\textrm{max, mass}$", "$J_\\textrm{max, area}$"
##   ),
##   format = "latex",
##   booktabs = TRUE,
##   escape = FALSE,
##   caption = "The number of non-missing (outside parentheses) and imputed (inside parentheses) observations of each trait for each plant functional type used in this analysis."
## ) %>%
##   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
##   group_rows("Broadleaf", 2, 6) %>%
##   group_rows("Needleleaf", 7, 9) %>%
##   group_rows("Shrub", 10, 12) %>%
##   group_rows("Grass", 13, 15) %>%
##   landscape()
```

```{r pairwise_correlation_table, echo = FALSE, results = "asis"}
rma_display <- readRDS(here::here("extdata", "rma-display.rds"))
kable(
  rma_display,
  "latex",
  booktabs = TRUE, longtable = TRUE,
  caption = paste("Pairwise reduced major axis (RMA) slope values",
                  "(Mean and 95pct credible interval)",
                  "for each plant functional type.",
                  "Values with asterisks indicate slopes",
                  "with credible intervals that don't intersect zero.")
)

## corr_results <- readd(corr_processed)
## kable(
##   corr_results %>%
##     dplyr::select(pft, yvar, xvar, corr_value, present, missing),
##   "latex",
##   booktabs = TRUE, longtable = TRUE, escape = TRUE,
##   col.names = c("PFT", "Trait 1", "Trait 2", "Correlation (95% CI)", "present", "missing"),
##   caption = paste(
##     "Pairwise trait correlation values (Mean and 95% credible interval)",
##     "for each plant functional type.",
##     "Values with asterisks indicate correlations",
##     "with credible intervals that do not intersect zero."
##   )
## )
```

\clearpage

# Method S1: Estimating trait means and covariances through multiple imputation {-}

## Introduction {-}

Under _single_ imputation, the imputation step occurs once outside of the MCMC loop.
In _multiple_ imputation, as implemented in this paper, the imputation step is part of the MCMC loop,
such that imputed data values are conditioned on the current state of the parameters and vice-versa.
In the figure below,
$Y$ is the original data (with missing values),
$Y^*$ is the original data with missing values imputed,
$\mu_p$ and $\Sigma_p$ are values of the mean vector and covariance matrix (respectively) calculated only from the original data,
and $\mu^*_t$ and $\Sigma^*_t$ are the draws of the mean vector and covariance matrix (respectively) at MCMC iteration $t$.

```{tikz diagram, echo = FALSE}
\usetikzlibrary{positioning, calc, fit}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 3) {Data $Y$};
\node (data params) at (-1, 2) {Calculate $\mu_p, \Sigma_p$};
\node (Ystar) at (-1, 1) {Impute $Y^* \mid \mu_p, \Sigma_p$};
\node (guess params) [align = center, right = of Ystar] {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (mu star) at (0, 0) {Draw $\mu^*_t \mid Y^*, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -1) {Draw $\Sigma^*_t \mid Y^*, \mu^*_t$};
\path [->] (Y) edge (data params);
\path [->] (data params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (guess params) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(mu star.east);
\node (mcmc loop) [fit = (mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Single imputation};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}[
	every node/.style = {shape = rectangle, draw = black},
	scale = 1.1
]
\node (Y) at (-1, 1) {Data $Y$};
\node (guess params) [align = center] at (1, 1) {Initial guess\\$\mu^*_0$ $\Sigma^*_0$};
\node (Ystar) at (0, 0) {Impute $Y^*_t \mid Y, \mu^*_{t-1}, \Sigma^*_{t-1}$};
\node (mu star) at (0, -1) {Draw $\mu^*_t \mid Y^*_t, \Sigma^*_{t-1}$};
\node (sigma star) at (0, -2) {Draw $\Sigma^*_t \mid Y^*_t, \mu^*_t$};
\path [->] (Y) edge (Ystar);
\path [->] (guess params) edge (Ystar);
\path [->] (Ystar) edge (mu star);
\path [->] (mu star) edge (sigma star);
\path [->, out = 0, in = 0] (sigma star.east) edge
	node (t plus one) [right, draw = none, font = {\footnotesize}] {$t = t + 1$}
	(Ystar.east);
\node (mcmc loop) [fit = (Ystar)(mu star)(sigma star)(t plus one), shape = rectangle, draw = black!50] {};
\node (mcmc loop text) [align = center, anchor = west, draw = none] at (mcmc loop.east) {MCMC\\loop};
\node (title) [anchor = north, draw = none] at (mcmc loop.south) {Multiple imputation};
\end{tikzpicture}
```

## Demonstration {-}

Consider two positively correlated traits, A and B.

```{r create_traits}
set.seed(12345) # For reproducibility
library(mvtraits)

true_mu <- c(0, 0)
true_cov <- matrix(c(1, 0.8, 0.8, 1), nrow = 2)
Y_all <- random_mvnorm(1000, true_mu, true_cov)
colnames(Y_all) <- c("A", "B")
plot(Y_all, pch = 19)
```

Simulate missingness by randomly removing half of the data.

```{r remove_na}
Y <- Y_all
Y[sample.int(2000, 1000)] <- NA
plot(Y_all, pch = 19, col = "grey")
points(Y, pch = 19, col = "black")
legend(
  "topleft",
  legend = c("True missing", "Present"),
  col = c("grey", "black"),
  pch = 19,
  bg = "white"
)
```

Set an uninformative multivariate normal prior on $\mu$...
```{r prior_mu}
mu_0 <- c(0, 0)
Sigma_0 <- diag(10, 2)
```

...and an uninformative Wishart prior on $\Sigma$.
```{r prior_sigma}
v0 <- 2
S0 <- diag(10, 2)
```

Take an initial guess at the mean and covariance by drawing from their priors.

```{r initial}
mu_star <- random_mvnorm(1, mu_0, Sigma_0)[1,]
mu_star
Sigma_star <- rWishart(1, v0, S0)[,,1]
Sigma_star
```

Impute values based on this initial guess.

```{r impute}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)

mylegend <- function() {
  legend(
    "topleft",
    legend = c("True missing", "Present", "Imputed"),
    col = c("grey", "black", "red"),
    pch = 19,
	bg = "white"
  )
}
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
mylegend()
```

Clearly, these initial values are a bad fit to the data.
Nevertheless, draw $\mu$ and $\Sigma$ conditioned on this set of imputed values.

```{r draw}
mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)
mu_star
Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
Sigma_star
```

Draw a new set of imputed values, conditioned on the current $\mu^*$ and $\Sigma^*$.

```{r impute_2}
Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
plot(Y_all, pch = 19, col = "grey")
points(Ystar, pch = 19, col = "red")
points(Y, pch = 19, col = "black")
mylegend()
```

This is still a bad fit to the data, but, though it is hard to tell, it has improved slightly.
Doing this repeatedly (in a loop) gradually improves the estimates of the covariance.

```{r loop}
par(mfrow = c(5, 4), mar = c(3, 3, 2, 0.1), cex = 0.4)
for (i in 1:20) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  title <- sprintf("iteration: %d, covariance: %.2f", i, Sigma_star[2, 1])
  plot(Y_all, pch = 19, col = "grey", xlab = "", ylab = "", main = title)
  points(Ystar, pch = 19, col = "red")
  points(Y, pch = 19, col = "black")
}
```

After a lot of MCMC samples, we can generate a distribution of covariance values, which provides an uncertainty estimate on the covariance.

```{r cov_estimate}
cov_ab <- numeric(1000)
for (i in seq_along(cov_ab)) {
  mu_star <- draw_mu(Ystar, Sigma_star, mu_0, Sigma_0)[1,]
  Sigma_star <- draw_Sigma(Ystar, mu_star, v0, S0)
  Ystar <- mvnorm_fill_missing(Y, mu_star, Sigma_star)
  cov_ab[i] <- Sigma_star[2, 1]
}
hist(cov_ab, xlab = "Covariance estimate", ylab = "Count", main = "Covariance estimate")
abline(v = true_cov[2, 1], col = "red", lwd = 2)
legend("topright", legend = "True cov.", lty = "solid", lwd = 2, col = "red")
```